---
title: "Sam Miller's Prosper Loan Data Wrangling"
output: html_document
---
This data set contains 113,937 loans with 81 variables on each loan, including 
loan amount, borrower rate (or interest rate), current loan status, 
borrower income, borrower employment status, borrower credit history, 
and the latest payment information.

I use this R file to wrangle the original data set (prosperLoanData.csv) into the values needed for my dimple.js/d3.js visualization


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE, warning=FALSE)
```

```{r, include = FALSE}
setwd(paste("/Users/samuelmiller/Documents/Udacity/Data\ Analyst\ Nanodegree/Data\ Visualization\ with\ D3/Final\ Project", sep=""))
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(lubridate)
library(gridExtra)
library(pander)
```

```{r}
Loans <- read.csv('prosperLoanData.csv', header = T, check.names = F, 
                  na.strings=c(""," ","NA"))
```

```{r}
Loans_Actual_Return <- Loans %>%
                          mutate(Rc_Complete = (LP_CustomerPayments - LoanOriginalAmount - LP_ServiceFees)/LoanOriginalAmount)%>%
                          mutate(Rc_Incomplete = (LP_CustomerPayments + LP_NonPrincipalRecoverypayments - LP_ServiceFees - LP_CollectionFees - LoanOriginalAmount - LP_NetPrincipalLoss)/LoanOriginalAmount)%>%
                          mutate(Ra_Complete = (((1+Rc_Complete)^(12))^(1/LoanMonthsSinceOrigination))-1)%>%
                          mutate(Ra_Incomplete = (((1+Rc_Incomplete)^(12))^(1/LoanMonthsSinceOrigination))-1)%>%
                          filter(EstimatedReturn != "NA")%>%
                          filter(ClosedDate != "NA")%>%
                          mutate(ActualReturn = ifelse( LoanStatus == "Chargedoff" | LoanStatus == "Defaulted", Ra_Incomplete, ifelse( LoanStatus == "Completed", Ra_Complete, NA)))

Loans$ListingCreationDateYear <- year(Loans$ListingCreationDate)
Loans_Actual_Return$LoanOriginationDateYear <- year(Loans_Actual_Return$LoanOriginationDate)

Loans_Actual_Return_Grouped <- Loans_Actual_Return%>%
                                    group_by(ListingCreationDateYear, ProsperRating.Alpha)%>%
                                    summarise(Total_Amount_of_Loans = n(), Median_Estimated_Return = median(EstimatedReturn), Median_Actual_Return = median(ActualReturn))

write.csv(Loans_Actual_Return_Grouped, "Prosper_Loans_Actual_Return_2.csv")
```